Fusing Filters with Integer Linear Programming

### 2014-05-12 2014betterfusionforfilters/Main.tex
Rewrote the abstract

\begin{abstract}
In this paper, we present an array fusion technique for functional programs, which is able to handle multi-loop programs containing size altering operations like filter. Recent work on array fusion shows how to extract data-flow graphs from functional programs and compile them into efficient imperative loops.  However, the compilation process only handles graphs that can be compiled into \emph{single} loops, and most programs require more. Other work demonstrated how use Integer Linear Programming (ILP) to \emph{cluster} the operators in a general data-flow graph into subgraphs that can be individually fused, but until now this approach did not handle operations like filter, which produce arrays of a different size than their input. The technique we introduce in this paper combines the advantages of both, by  extending the existing ILP approach with support for size changing operators, using an external ILP solver to find good clusterings.

>>>>

\begin{abstract}
Minimising memory traffic is key to compiling functional, collection oriented array programs into efficient code. To achieve this, it is not sufficient to just fuse subsequent array operations into a single, more complex computation, we also have to also \emph{cluster} separate traversals of the same array into one single loop. Previous work demonstrated how Integer Linear Programming (ILP) can be used to cluster the operators in a general data-flow graph into subgraphs, which can be individually fused. However, these approaches have one serious restrictions: they can only handle operations which preserve the size of the array. The technique we present in this paper extends the ILP approach with support for size-changing operations, using an external ILP server to find good clusterings.
